Manuscript 1 code

### 2024-09-24 manuscript_1_code.Rmd
added 'mytheme' code to shorten some of the ggplots :tropical_fish:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
```

@@

## Figure formatting
```{r figure formatting info}
habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

@@

## plot
```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  # scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

```{r Figure 1 v2}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = unique(figure_1_data$year)) +
    labs(y = "Percent Cover",
         x = "Year",
         title = hab) +  # Add habitat as title
    theme_bw() +
    theme(axis.title.x = element_text(size = 18),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
          axis.title.y = element_text(size = 18),
          axis.text.y = element_text(size = 12),
          plot.title = element_text(size = 20, hjust = 0),
          legend.text = element_text(size = 14), 
          legend.title = element_blank(),
          legend.position = 'bottom',
          legend.box = 'horizontal',
          legend.box.background = element_rect(colour = "black"))
})

@@

### plot
```{r figure 2}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "Quadrat-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1, size=14),
        axis.text.y = element_text(size=12),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

@@

### plot
```{r}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
  geom_jitter(data = alpha_diversity_site_macro, aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 14, yend = 14, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "Site-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1, size=14),
        axis.title.y = element_text(size=18), 
        axis.text.y = element_text(size=14),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(dplyr)
```

@@

## Figure formatting
```{r figure formatting info}
mytheme <- theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))

habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

@@

## plot
```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  # scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  mytheme)

@@

## plot after aes changes
```{r Figure 1 v2}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = unique(figure_1_data$year)) +
    labs(y = "Percent Cover",
         x = "Year",
         title = hab) +  # Add habitat as title
    theme_bw() +
    mytheme
})

@@

### plot
```{r figure 2}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "Quadrat-level Richness") +
  theme_bw() + 
  mytheme +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=12),
        plot.title = element_text(size=18)))

@@

### plot
```{r}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
  geom_jitter(data = alpha_diversity_site_macro, aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 14, yend = 14, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "Site-level Richness") +
  mytheme +
   theme(axis.text.x = element_text(size=14),
      axis.text.y = element_text(size=14),
      plot.title = element_text(size=18),
      ))
